EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-03-27 edipack2.tex
Updated draft, up to Lanczos diagonalizaion. Let's speed up the writing huh...

\usepackage{listings}
\lstdefinestyle{fstyle}{
  frame=lines,
  language=fortran,
  basicstyle=\footnotesize,
  stringstyle=\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\color{magenta},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt
}

@@

\title{EDIpack2: interoperable Lanczos-based exact diagonalization solver for
  generic quantum impurity problems.}
% \author[]{L.~Crippa}
% \author[]{I.~Krivenko}
% \author[]{S.~Giuli}
% \author[]{G.~Bellomia}
% \author[]{....}
% % \author[]{A.~Scazzola}
% % \author[]{F.~Petocchi}
% % \author[]{G.~Mazza}
% % \author[]{L.~de Medici}
% % \author[]{M.~Capone}
% \author[a]{A.~Amaricci\corref{author}}

% \cortext[author] {Corresponding author.\\\textit{E-mail address:} amaricci@iom.cnr.it}

% \address[a]{CNR-IOM, Istituto Officina dei Materiali,
%     Consiglio Nazionale delle Ricerche, Via Bonomea 265, 34136 Trieste, Italy}
\begin{abstract}

@@

\section{Installation}\label{SecInstall}
The installation of \NAME is available through CMake which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries: {\tt
  libedipack.a} and {\tt libedipack\_cbinding.so}.
The former, alongside the generated Fortan modules files,
wraps the \NAME software. The latter enables interoperability through
bindings to the C language.

\subsection{Dependencies}
Although \NAME is as much self-contained as possible its development
hinges on two external libraries.
\begin{itemize}
\item {\bf SciFortran}: an open-source Fortran library to support
  mathematical and scientific software development. 
\item {\bf MPI} (optional): a distributed memory parallel communication layer with support to modern Fortran compiler.
\end{itemize}

@@

\subsection{Build and Install}
\subsubsection{Source}
The software can be installed from source as follows. The source can
be retrieved directly from its GitHub repository, for instance using:
\begin{lstlisting}[language=Bash]
git clone https://github.com/edipack/EDIpack2.0 EDIpack2
\end{lstlisting}
Then, assuming to be in the software directoru, a conventional
out-of-source building is performed using two different compilations
backends.

\begin{itemize}
  \item {\bf GNU Make}\\
This is the default CMake workflow:
\begin{lstlisting}[language=Bash]
mkdir build
cd build
cmake ..
make -j
make install
\end{lstlisting}

@@

An alternative workflow employs the Ninja building backend with
Fortran support. Ninja is generally faster and automatically supports
multi-threaded building:
\begin{lstlisting}[language=Bash]
mkdir build
cd build
cmake -GNinja ..
ninja
ninja install
\end{lstlisting}
\end{itemize}

@@

The Conda package installation procedure reads:
\begin{lstlisting}[language=Bash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}
\noindent
which installs a bundle of Scifor and \NAME libraries together with
specific {\tt pkg-config} configurations files which can be used to
retrieve compilation and linking flags.

@@

\begin{lstlisting}[language=Bash]
git clone https://github.com/edipack/EDIpy2 EDIpy2
cd EDIpy2
pip install . 
\end{lstlisting}
In some more recent Python distribution the flag {\tt
  --break-system-packages} might be required to complete
installation or a virtual environment should be used instead.

@@

\begin{lstlisting}[language=Bash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}

@@

\begin{lstlisting}[language=Bash]
git clone https://github.com/krivenko/edipack2triqs
cd edipack2triqs
pip install .
\end{lstlisting}

@@

@@

\subsection{The quantum impurity problem}\label{sSecQIM}

\subsection{The  basis states}\label{sSecBasis}

\subsection{Conserved quantum numbers}\label{sSecQNs}

>>>>

\usepackage{listings}
\lstdefinestyle{fstyle}{
  frame=lines,
  language=fortran,
  basicstyle=\footnotesize,
  stringstyle=\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\color{red},
  % morekeywords={*,...},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt
}

\lstdefinestyle{mybash}{
  frame=lines,
  language=bash,
  basicstyle=\footnotesize,
  captionpos=b,
  stringstyle=\color{mymauve},%\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\bfseries,
  morekeywords={*,git,mkdir,cmake,make},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt
}

% % MINTED:
% % \usepackage[cache=true,cachedir=minted-cache]{minted}
% % \usepackage[finalizecache=true,cachedir=minted-cache]{minted}
% \usepackage[frozencache=true,cachedir=minted-cache]{minted}
% \setminted[fortran]{linenos,mathescape,frame=lines,style=bw,framesep=1mm,
%   baselinestretch=1,fontsize=\footnotesize}

@@

\title{EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems}
\author[a,b,c]{L.~Crippa\corref{AA}}
\author[a]{I.~Krivenko}
\author[d]{S.~Giuli\corref{BB}}
\author[d]{G.~Bellomia\corref{CC}}
\author[e]{A.~Scazzola}
\author[f]{F.~Petocchi}
\author[g]{G.~Mazza}
\author[h]{L.~de Medici}
\author[d]{M.~Capone}
\author[i]{A.~Amaricci}

% \cortext[author] {Corresponding author.\\\textit{E-mail address:} amaricci@iom.cnr.it}
\newcommand{\CNRIOM}{CNR-IOM, Istituto Officina dei Materiali,
  Consiglio Nazionale delle Ricerche, Via Bonomea 265, 34136
  Trieste, Italy}
\newcommand{\SISSA}{Scuola Internazionale Superiore di Studi Avanzati (SISSA),
  Via Bonomea 265, 34136 Trieste, Italy}
\newcommand{\ITPHamburg}{I. Institute of Theoretical Physics,
  University of Hamburg, Notkestrasse 9, 22607 Hamburg, Germany}
\newcommand{\WZBURG}{Institut f\"ur Theoretische Physik und
  Astrophysik,Universit\"at W\"urzburg, 97074 W\"urzburg, Germany}
\newcommand{\CTQMAT}{W\"urzburg-Dresden Cluster of Excellence ct.qmat, 01062 Dresden, Germany}
\newcommand{\Geneve}{Department of Quantum Matter Physics, University of
  Geneva, Quai Ernest-Ansermet 24, 1211 Geneva, Switzerland}
\newcommand{\UPISA}{Department of Physics ``E. Fermi'' University of
  Pisa, Largo B. Pontecorvo 3, 56127 Pisa, Italy}
\newcommand{\ESPCI}{LPEM, ESPCI Paris, PSL Research University, CNRS, Sorbonne Universit\'e, 75005 Paris, France}

\address[a]{\ITPHamburg}
\address[b]{\CTQMAT}
\address[c]{\WZBURG}
\address[d]{\SISSA}
\address[e]{Torino}
\address[f]{\Geneve}
\address[g]{\UPISA}
\address[h]{\ESPCI}
\address[i]{\CNRIOM}
\cortext[AA] {Corresponding author \textit{E-mail address:}
  lorenzo.crippa@uni-hamburg.de}
\cortext[BB] {Corresponding author \textit{E-mail address:}
  giuli@sissa.it}
\cortext[CC] {Corresponding author \textit{E-mail address:}
  bellomia@sissa.it}

\begin{abstract}

@@

\section{Installation}\label{SecInstall}
The installation of \NAME is available through CMake which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries.
The main one is {\tt  libedipack.a} which, alongside the generated Fortan
modules, wraps the \NAME software possibly including support for  
inequivalent impurities.
A second dynamic library, {\tt libedipack\_cbinding.so}, enables interoperability through
specific bindings to the C programming language.

\subsection{Dependencies}
\NAME essentially depends on two external libraries.
\begin{itemize}
\item {\bf SciFortran}: an open-source Fortran library to support
  mathematical and scientific software development. 
\item {\bf MPI} (optional): a distributed memory parallel communication layer with support to modern Fortran compiler.
\end{itemize}

@@

\subsection{Build and Install}
\subsubsection{Source}
The software can be installed from source as follows. The source can
be retrieved directly from its GitHub repository, for instance using:
\begin{lstlisting}[style=mybash]
git clone https://github.com/edipack/EDIpack2.0 EDIpack2
\end{lstlisting}
Then, assuming to be in the software directoru, a conventional
out-of-source building is performed using two different compilations
backends.

\begin{itemize}
  \item {\bf GNU Make}\\
This is the default CMake workflow:
\begin{lstlisting}[style=mybash]
mkdir build
cd build
cmake ..
make -j
make install
\end{lstlisting}

@@

An alternative workflow employs the Ninja building backend with
Fortran support. Ninja is generally faster and automatically supports
multi-threaded building:
\begin{lstlisting}[style=mybash]
mkdir build
cd build
cmake -GNinja ..
ninja
ninja install
\end{lstlisting}
\end{itemize}

@@

The Conda package installation procedure reads:
\begin{lstlisting}[style=mybash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}
\noindent
which installs a bundle of Scifor and \NAME libraries together with
specific {\tt pkg-config} configurations files which can be used to
retrieve compilation and linking flags.

@@

\begin{lstlisting}[style=mybash]
git clone https://github.com/edipack/EDIpy2 EDIpy2
cd EDIpy2
pip install . 
\end{lstlisting}
In some more recent Python distribution the flag {\tt
  --break-system-packages} might be required to complete
installation or a virtual environment should be used instead.

@@

\begin{lstlisting}[style=mybash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}

@@

\begin{lstlisting}[style=mybash]
git clone https://github.com/krivenko/edipack2triqs
cd edipack2triqs
pip install .
\end{lstlisting}

@@

% $H_{e-ph}= \sum_{\a\sigma} g_\a
% d^+_{\a\sigma}d_{\a\sigma}(b+b^+)$.

@@

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a general quantum impurity problem defined by the
following Hamiltonian:
$$
\hat{H} = \hat{H}_{imp} + \hat{H}_{bath} + \hat{H}_{hyb} + \hat{H}_{ph} + \hat{H}_{e-ph}
$$
which describes a  multi-orbital interacting quantum
impurity coupled to an electronic bath and to local, i.e. Holstein,
phonons. We assuming for the moment that no particular symmetry
holds. The impurity part of the Hamiltonian reads: 
\begin{equation}\label{Himp}
  \begin{split}
    \hat{H}_{imp} & = \hat{H}^0_{imp} + \hat{H}^{int}_{imp}\\
    \hat{H}^0_{imp} & =
    \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}d_{\b\sigma'}\\
    %
    \hat{H}^{int}_{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}\\
    % 
\end{split}
\end{equation}

where $d_{\a\sigma}$ ($d^+_{\a\sigma}$) are the destruction (creation)
second-quantization operators for impurity electrons in the
orbital $\alpha=1,\dots,N_\a$, with $N_\a$ the number of orbitals,
with spin $\sigma=\up,\dw$ and whose occupation is described
by the  operator  
$n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$. 
The non-interacting internal structure of the impurity is described by
the $h^{0}_{\a\b\sigma\sigma'}$ matrix. 
$\hat{H}^{int}$ describes the  local multi-orbital
interaction~\cite{Georges2013ACMP} which, for simplicity, we take as a
generalized Hubbard-Kanamori form.  
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The  remaining two terms are, respectively, the spin-exchange and
pair-hopping which we considered with their respective independent
couplings $J_X$ and $J_P$.
In the case $N_\a=3$ a fully symmetric $SU(3)_{orbital}\times SU(2)_{spin}\times
U(1)_{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other values of
$N_\a$~\cite{Georges2013ACMP}.

The bath part and its coupling to the impurity has the form: 
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^{+}_{p\a\sigma}a_{p\b\sigma'}\\
    %
    \hat{H}_{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}a_{p\b\sigma'}+H.c. \\
\end{split}
\end{equation}
where $p=1,\dots,N_{bath}$ is an index running over a finite number of
bath elements, $a_{p\alpha\sigma}$ ($a^+_{p\alpha\sigma}$) are the destruction (creation) operators for
the bath electrons with index $p$, with orbital $\alpha$ and
spin $\sigma$.
The  properties of each bath level are described by the
matrices $h^p_{\a\b\sigma\sigma'}$. As such any bath element can be
composed of several electronic levels according the bath topology,
which will be discussed further in the following.
Each bath level couples to the impurity with an amplitude
$V^p_{\a\b\s\sigma'}$ which we allow to couple different orbital and
opposite spins.

Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, $b_q$
($b_q^+$) are the destruction (creation) operators for the phonon $q$
with frequency $\omega_{0q}$. The matrix  $g_{\a\b}$ expresses is the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$.

In the following we consider a bath discretized into a  number
of bath degrees of freedom and a finite number of available phonons,
to cut-off the unbounded dimensions of the local phonons Hilbert space.

More specifically, we consider a system composed of $N_{imp}=1$
impurities, i.e. a single impurity problem, $N_{bath}$ bath elements
and $N_{ph}$ phonons. The size of the system is determined by
the number of phonons (fixed) and that of {\it electronic} levels, i.e. levels with a local
electronic Hilbert space
$\HH_e=\{\ket{0},\ket{\up},\ket{\dw},\ket{\up\dw} \}$. 
Due to is internal structure, the single impurity contains
$N_i= N_\a$ electronic levels.
The total number of levels is then determined by
the electronic levels in the bath $N_b$. This is a function of the bath topology and $N_{bath}$,
i.e. the number of bath elements. In the simplest case each bath
element corresponds to an independent electronic levels coupled to the
impurity, thus $N_b\equiv N_{bath}$. 
We indicate with $N_s=N_i + N_b$ the total number of electronic levels.

\subsection{The Fock basis states}\label{sSecBasis}
The Fock space of the quantum impurity problems is defined as
$\FF=\FF_e\otimes \FF_{ph}$, with $\FF_e=\bigoplus_{n=0}^{N_s}
S_-\HH_e^{\otimes n}$ the electronic Fock space,  $\FF_{ph}=\bigoplus_{n=0}^{N_q}S_+\HH_{ph}^{\otimes n}$ the
phonon Fock space,  $\HH_{ph}=\{\ket{0},\ket{1},\dots,\ket{N_{ph}}\}$ is the local phonon Hilbert space
and ($S_-$)  $S_+$ the (anti-)symmetrization operator.  
The total dimension of the Fock space is
$D=D_e\cdot D_{ph}=4^{N_s}\cdot (N_{ph}+1)$ making the exponential
growth with the number of electron levels transparent. 
The quantum states in the space $\FF$ are naturally represented in
terms of occupation number formalism of the second quantization,
i.e. the Fock basis. For a system of $N_s$ electrons each Fock state
reads $\ket{p}\ket{\vec{n}}$ with $\ket{\vec{n}}=\ket{\vec{n}_\up}\ket{\vec{n}_\dw}=\ket{n_{1\up},\dots,n_{N_s\up},n_{1\dw},\dots,n_{N_s\dw}}$, 
where $p=1,\dots,N_{ph}$ is the number of local phonons while $n_{a\sigma}=0,1$ signals the absence or the
presence of an electron with spin $\sigma$ at the level $a$.
Remarkably, the electronic part of the Fock state $\ket{\vec{n}}$ is represented as a string of
zeros and ones of length $2N_s$. Thus, any such state can  be encoded
in a computer using a sequence of $2N_s$ bits or, analogously, as a
given integer $I=0,\dots 2^{2N_s}-1$ so that $\ket{\vec{n}}=\ket{I}$.  
Together with the basis states one defines  destruction and creation 
operators, respectively $c_{a\s}$ and $c^+_{a\s}$, which acts on the
Fock space as: 
$\ket{\vec{n}}$ as:  
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\qquad
    c^{+}_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ takes care
of the fermionic sign imposed by Pauli principle.

\subsection{Conserved quantum numbers}\label{sSecQNs}
In order to circumvent the exponential scaling of the dimensions of
the problem scales it is necessary to take into account suitable symmetries.
Indeed, considering operators ${\cal Q}$ such that $[H, {\cal Q}]=0$
reduces of the Fock space into several symmetry sectors
labelled by given quantum numbers $\vec{Q}$. 
In the context of quantum impurity problems two symmetries are often considered: i)
conservation of the total occupation $N$ and ii) the
conservation of the total magnetization $S_z$. Although the total spin
operator $S^2$ can also be conserved, the difficult  implementation
and the marginal gain makes it often non convenient to include this symmetry.

In \NAME we consider three different cases:
\begin{itemize}
\item{\bf normal:}  $\vec{Q}=[N,S_z]\equiv[N_\up,N_\dw]$
\item{\bf superc:}  $\vec{Q}=S_z$
\item{\bf nonsu2:} $\vec{Q}=N$
\end{itemize}

The {\bf normal} case consider independent conservation of the total occupation $N$
and the total magnetization $S_z$ or, equivalently, the total number of electrons with spin up $N_\up$ and down
$N_\dw$. Optionally, we consider the case in which the symmetry
applies separately for each orbital and spin,
i.e. $\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This special case has been
extensively discussed in \onlinecite{Amaricci2022} so it won't be covered in
this work.

The {\bf superc} case deals with the conservation of the total
magnetization only, so that total charge may not be conserved. This
case includes a description of $s$-wave superconductivity, featuring
intra- and inter-orbital components.

Finally, the {\bf nonsu2} case consider the conservation of the total
charge, whereas the spin symmetry group is not fully
conserved. Although there many possible realization of this scenario,
this particularly applies to the presence of local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}, the emergence of in-plane spin ordering~\cite{KM} or
in-plane spin-triplet exciton condensation~\cite{ExcitonPRB,Amaricci,Blason}.

From a computational point of view the construction of a symmetry
sector corresponds to the determination of a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF$ relating the states $\ket{i}$
belonging to the sector $\SS_{\vec{Q}}$ to the states $\ket{I}$ in the
Fock space. Operatively, the map corresponds to an integer rank-1
array of dimension $D_\SS$ which is the {\it dimension} of the
sector. According to the previously introduced symmetries we
have:
\begin{itemize}
\item{\bf normal:}  $D_\SS = \binom{N_s}{N_\up}\binom{N_s}{N_\dw}$ 
\item{\bf superc:}  $D_\SS = \sum_m 2^{N_s-S_z-2m}\binom{N_s}{N_s-S_z-2m}\binom{S_z+2m}{m}$
\item{\bf nonsu2:} $D_\SS=\binom{2N_s}{N}$
\end{itemize}

The {\bf normal} case requires a brief remark. Because of the
independent conservation of $N_\up$ and $N_\dw$, the sector can be further factorized into $\SS_{\vec{Q}} = \SS_{N_\up}\otimes
\SS_{N_\dw}$ and, correspondingly, the sector map reads $\MM = \MM_\up
\otimes \MM_\dw$.
As such, each state $\ket{i}=\ket{i_\up}\ket{i_\dw}$ of the
sector is labelled by two integers $[i_\up,i_\dw]$, 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$. The map $\MM$ connects any such state to a Fock state
$\ket{I}=\ket{I_\up}\ket{I_\dw}$ labelled by two integers
$[I_\up,I_\dw]$ as $I=I_\up +   I_\dw 2^{N_s}$. For a thorough
discussion about Fock basis organization in this case see
\onlinecite{amaricci2022}.

The symmetry sector reduction of the Fock space induces a block
digonal form to the Hamiltonian matrix. Each block, labelled by a
value of the quantum numbers $\vec{Q}$ has dimension
$D_{\SS(\vec{Q}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix. In the {\bf
  normal} case $H_\SS$ takes a particularly symmetric form thanks to
the product structure of the sector and its map~\cite{amaricci2022}.
The analysis of the spectrum is then reconducted to the inspection of
the Hamiltonian in each symmetry sector. Should particular constraint
holds the search can be limited to only particular sector, further
reducing the computational cost.

Although the sectors have dimensions much smaller than the full Fock
space, for large systems storing the Hamiltonian matrix in the memory
can still be highly inefficient.
In such cases, Krylov or Lanczos methods~\cite{Lanczos1950JRNBSB,Lin1993CIP,Lehoucq1998,Maschhoff1996} can
be implemented using a storage-free algorithm, performing the
necessary linear operations on-the-fly.
This solution has generally a negative impact on the execution
time, however this can be well compensated by scaling in a distributed 
parallel framework.
